---
title: 'Projects and GitHub'
output:
  html_document: default
  pdf_document:
    fig_height: 3
    fig_width: 4.5
word_document: default
---
  
```{r, setup, include=FALSE}
library(tidyverse)
```

##R projects, project sharing, and version control (GitHub)

###R Projects

Want to manage your R work like an expert?  R experts keep all the files associated with a project together — input data, R scripts, analytical results, figures. This is such a wise and common practice that RStudio has built-in support for this via **projects**.

> *Beware*: when you complete the steps below, you will clear all the variables and data sets in your environment, all the plots under Plots, all the libraries you've loaded, and all the output in your console.  Actually, this is a good thing.  All of these things are temporary, while your R scripts and Rmd files are "real".  In fact, R4DS recommends never preserving your workspace between session (under Tools > Global Options) -- it might cause some short term pain, but it forces you to capture important interactions in your code.

Let's create a project for this web scraping work.  You can start in the upper left under File > New Project, or in the upper right under Project: (None) > New Project.  Select New Directory > Empty Project.  Enter a directory name (like Scraping), make sure it's being placed in the right subdirectory (like your student folder) and hit Create Project.  You should now have a blank slate.

First move imdbScrape.Rmd into your Scraping folder or make a copy in your Scraping folder.  Knit the file.  Note that there is still nothing in the Environment; when you knit, everything happens within that process.  Now go back into imdbScrape.Rmd and run the first few chunks by hand, until `omdb` is generated.  Then copy the line below into imdbScrape.Rmd (or simply copy this file into the Scraping folder) and run it to write `omdb` to your project folder:

```{r}
write_csv(omdb, "omdb.csv")
```

Not only should there be items in the Environment now, but `omdb.csv` should have been written to your Scraping folder.  To see the power of projects, select File > Quit Session once again, and this time select Scraping under Projects.  You will get a clean Environment and Console once again, but History shows the commands you ran by hand, active Rmd files appear in the Source panel, and Files contains the Rmd, html, and csv files produced by your last session.

###Project Sharing

Now let's see another advantage of using R projects - being able to share materials with collaborators.  To do this, go to File > Share Project.  Type a collaborators username and hit Add.  Your collaborator can work on your project by using the URL listed or simply by finding the project folder in their Home directory.  A project collaborator has the same permissions that you do on all project files, so you must coordinate efforts.  In addition, you can all see each other perform real-time editing, but anyone can save all changes at any time, so again you must coordinate.

For more information, see: https://wp.stolaf.edu/it/project-sharing/

###Version Control (GitHub)

Another way to collaborate on an R project (or any coding project in general) is by using a version control system like GitHub.  With GitHub, you never have to save files as Final.docx, Final2.docx, Newfinal.docx, RealFinal.docx, nothisisreallyit.docx, etc.  You update code like .R and .Rmd and data files, recording descriptions of any changes.  Then if you ever want to go back to an earlier version, GitHub can facilitate that.  Or if you want to make your work public, others can see it and even suggest changes, but you are ultimately in control of any changes that get made.  And, like R projects, you can have multiple collaborators with access to the same set of files.

